A Graph Theory Based Approach to MDAO Problem Formulation

### 2013-04-13 graph_syntax.tex
cleaning up merge conflict stuff some more

\subsection{Design Variables}
((DP:I'm thinking we might also need the upper indegree limit to be zero as well, because setting only the lower indegree limit makes the node a design variable if it didn't have any incoming edges. It's seems possible that in the MCG a node has incoming edges but the user wants it to be a design variable. I can make the change if you agree.))
For the Sellar problem, there are three design variables, $x_1^*$, $z_1^*$, and $z_2^*$. In Fig. 
\ref{f:sellar_graph_full}, these are the only three holes (nodes without out 
incomming edges) in the graph. Based on this example it's tempting to simply 
call all holes in the graph design variables. However, in the general case this is 
not completely correct. For any given problem, the set of holes only identifies 
all the potential design variables. However, some variable holes might not make good design 
variables. For instance, if your aircraft mission analysis code requires a drag input 
which ends up being a hole then it's likely that some new analysis tool is needed
to predict drag before the design process can continue.

@@

More formally stated, given a MCG with the $\txt{deg}_l^-(v)=1$ for all nodes, 
there will be a set of holes. Each hole represents a node that could potentially 
become a design variable. It is up to the designer to examine that variable and 
decide if it is appropriate to use it as a design variable. If a variable should 
be a design variable, then set it's ${deg}_l^-(v)=0$. Otherwise you leave 
${deg}_l^-(v)=1$ and find some tool to fill the hole in graph.

So a design variable is defined as any variable node with ${deg}_l^-(v)=0$. 
This, by definition, means that design variables are not holes in the graph, since 
they have not violated their lower in-degree limit. Regardless, 
when working with a MCG or FPG, only connection edges are allowed, so a design 
variable in either graph will be a variable with no incomming edges edges. For a valid PSG 
any design variable node would have at least one, but possibly multiple, incomming 
implicit edges from one or more driver nodes.

@@

  A simple definition of a global node is any node that involves information 
  from more than one analysis block. Likewise, a local node is any node 
  that involves information from only a single analysis block. As stated before, variable 
  nodes with fixed edges directed to model nodes (as part of an analysis block) are inherently local 
  and similarly model nodes are also local. Expression and driver nodes
  are not inherently local or global though. For these, their locality is determined 
  by the incomming and outgoing edges they possess. Consider the graph in Fig. \ref{f:local_global}, 
  where the expression node $g_1$ represents some kind of constraint for the optimziation problem. 
  Since $g_1$ only has edges comming from analysis block $B$, it would be considered a local constraint 
  on block $B$. $g_2$ has incomming edges from both $A$ and $B$ analysis blocks, so it would be a global constraint.

>>>>

\subsection{Design Variables}
((DP:I'm thinking we might also need the upper indegree limit to be zero as well, because setting only the lower indegree limit makes the node a design variable if it didn't have any incoming edges. It's seems possible that in the MCG a node has incoming edges but the user wants it to be a design variable. I can make the change if you agree.))
For the Sellar problem, there are three design variables, $x_1^*$, $z_1^*$, and $z_2^*$. In Fig. 
\ref{f:sellar_graph_full}, these are the only three holes (nodes without out 
incomming edges) in the graph.

@@

More formally stated, given a MCG with the $\txt{deg}_l^-(v)=1$ for all nodes, 
there will likely be a set of holes preventing a valid FPG from being obtained. 
Each hole represents a node that could potentially become a design variable. 
It is up to the designer to examine each and decide if it is appropriate to 
consider it a design variable, in which case the designer shall set the lower 
indegree limit to be zero. However, some variable holes may not actually be 
suitable as a design variable. For instance, if an aircraft mission analysis 
code requires a drag input which ends up being a hole then it's likely best to 
leave$\txt{deg}_l^-(v)=1$ and find a new analysis code to fill the hole in graph.

So a design variable is defined as any variable node with $\txt{deg}_l^-(v)=0$. 
This, by definition, means that design variables are not holes in the graph, since 
they have not violated their lower in-degree limit. Regardless, 
when working with a MCG or FPG, only connection edges are allowed, so a design 
variable in either graph will be a variable with no incomming edges edges. For a valid PSG 
any design variable node would have at least one, but possibly multiple, incomming 
implicit edges from one or more driver nodes.

@@

  A simple definition of a global node is any node that involves information 
  from more than one analysis block. Likewise, a local node is any node 
  that involves information from only a single analysis block. As stated before, variable 
  nodes with fixed edges directed to model nodes (as part of an analysis block) are inherently local 
  and similarly model nodes are also local. Global variables are a part of many 
  MDAO problems and must be representable in the graph. Follwoing our simple definition
  any variable node that had multiple incomming or outgoing edges, would be global. 
  In Fig. \ref{f:sellar_graph_full} the variable nodes for $x_1^*$, $z_1^*$, and $z_2^*$ 
  have multiple outgoing edges and would be considered global.

  Note that the variable $x_1$ from the Sellar problem is usually refered to as 
  a local variable, which is contradicted by the given graph. This discrepancy 
  arises from the explicit treatment of objectives and constraints as separate 
  expression blocks. Normally in the Sellar problem $x_1$ is considered a local 
  variable because it only directly affects analysis block $D_1$. By forcing
  the expansion of $F$ into an expression block with its own inputs, then $F$ 
  gets it's own copy of the $x_1$ variable. This necessitates the creation 
  of the global $x_1^*$ node in the graph with two outgoing edges to link the 
  two $x_1$ inputs nodes in the different blocks. Interestingly, using this 
  definition, where $x_1^*$ is a global variable fits nicely within 
  the structure of Braun's Colaborative Optimization (CO) archtiecture \cite{braun_thesis}. 
  The rules of setting up a problem with CO require that each discipline be 
  given an independent local copy of all global variables, while any local 
  variables are retained uniquely within their respective disciplines. But 
  when a local design variable shows up explicitely in the objective function 
  or global constraint then a global variable is created, and the discipline again 
  gets an independent local copy of it. Given our definition of global vs local
  variable nodes, no such exception for local variables is necessary. When you 
  include a given variable in an expression node, it becomes global by definition 
  and should be treated as such in CO.

  Analysis and expression blocks can also have a locality defined for the set of 
  nodes and edges within the block. Their locality is determined by the 
  incomming and outgoing edges from the block as a whole. Fig. \ref{f:sellar_graph_full} has expression blocks for each constraint, $G_1$ and $G_2$. Both constraints are 
  local to their respective analysis blocks. But the expression block for $F$ 
  has incomming edges from $D_1$ and $D_2$, so it would considered global.
